---
title: "Correlation?"
output:
  html_document:
    #code_folding: hide
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
  html_notebook:
    #code_folding: hide
    toc: yes
---
```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```
## Preparing data for analysis 
First thing what I need to do is to count how many customers align additional agreement and calculate the total for specific loan. 

```{r}
#Additional.Amount by loan number
subset_AA <- aggregate(test$Additional.Amount, by=list(Category=test$loan.number), FUN=sum)
names(subset_AA)[1] <- "loan.number"
names(subset_AA)[2] <- "Additional.Amount_SUM"
cat('Number of rows subset_AA: ', nrow(subset_AA), ' ')

#calculate number of additional aggreement by loan
subset_AA_ <- aggregate(test$loan.number, by=list(Category=test$loan.number), FUN=length)
names(subset_AA_)[1] <- "loan.number"
names(subset_AA_)[2] <- "Number_OF_AA"
cat('Number of rows subset_AA_: ', nrow(subset_AA_))
#merge
subset_AA <- merge(x = subset_AA, y = subset_AA_, by = "loan.number", all = TRUE)

head(subset_AA)
```
Let's join data from subset_AA with MA and replace all NA with 0

```{r}
#check if loan.numbr unique
anyDuplicated(MA$loan.number)
Join_aggr <- merge(x = MA, y = subset_AA, by = "loan.number", all = TRUE)
head(Join_aggr)

#let's connect the rest of transposes data
Join_aggr <- merge(x = Join_aggr, y = test_2, by = "loan.number", all = TRUE)
head(Join_aggr)

#replace 0 instead NA
Join_aggr <- as.data.frame(Join_aggr)
Join_aggr$Additional.Amount_SUM[is.na(Join_aggr$Additional.Amount_SUM)] <- 0
Join_aggr$Number_OF_AA[is.na(Join_aggr$Number_OF_AA)] <- 0
Join_aggr$Additional.Amount_1[is.na(Join_aggr$Additional.Amount_1)] <- 0

Join_aggr$Additional.Amount_2[is.na(Join_aggr$Additional.Amount_2)] <- 0
Join_aggr$Additional.Amount_3[is.na(Join_aggr$Additional.Amount_3)] <- 0
Join_aggr$Additional.Amount_4[is.na(Join_aggr$Additional.Amount_4)] <- 0
Join_aggr$Additional.Amount_5[is.na(Join_aggr$Additional.Amount_5)] <- 0
Join_aggr$Additional.Amount_6[is.na(Join_aggr$Additional.Amount_6)] <- 0
Join_aggr$Additional.Amount_7[is.na(Join_aggr$Additional.Amount_7)] <- 0
Join_aggr$Additional.Amount_8[is.na(Join_aggr$Additional.Amount_8)] <- 0
Join_aggr$Additional.Amount_9[is.na(Join_aggr$Additional.Amount_9)] <- 0
Join_aggr$Additional.Amount_10[is.na(Join_aggr$Additional.Amount_10)] <- 0
Join_aggr$Additional.Amount_11[is.na(Join_aggr$Additional.Amount_11)] <- 0
Join_aggr$Additional.Amount_12[is.na(Join_aggr$Additional.Amount_12)] <- 0
Join_aggr$Additional.Amount_13[is.na(Join_aggr$Additional.Amount_13)] <- 0
Join_aggr$Additional.Amount_14[is.na(Join_aggr$Additional.Amount_14)] <- 0


Join_aggr$`Loan status` <- as.numeric(Join_aggr$`Loan status`)
Join_aggr[,c(1,2,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25)]  %>% replace(is.na(.), 0)

#head(Join_aggr)

```

## Correlation 

Let’s examine what the Pearson correlation shows to us
```{r message=FALSE, warning=FALSE}
#cor(Join_aggr[,c(1,2,4,5,6,7,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25)])
cor(x[,unlist(lapply(MA, is.numeric))])
pairs.panels(MA[1:7])  


```

Let's includes significance indicators

```{r message=FALSE, warning=FALSE}
chart.Correlation(x[2:11])

```

```{r} 
corr_pearson <- cor.test(x$`Extensions per Loan`, x$Additional.Amount_SUM, 
                    method = "pearson")
corr_pearson

corr_spearman <- cor.test(x$`Extensions per Loan`, x$Additional.Amount_SUM, 
                    method = "spearman")
corr_spearman
```



```{r}
#hist(Join_aggr$`Closed Date`,breaks=10)
#hist(Join_aggr$`Signed Date`,breaks=10)
#format the date to month
Join_aggr$`Closed Date` <- format(Join_aggr$`Closed Date`,"%m")
Join_aggr$`Closed Date` <- as.numeric(Join_aggr$`Closed Date`)
Join_aggr$`Signed Date` <- format(Join_aggr$`Signed Date`,"%m")
Join_aggr$`Signed Date` <- as.numeric(Join_aggr$`Signed Date`)

Join_aggr$created_1 <- format(Join_aggr$created_1,"%m")
Join_aggr$created_2 <- format(Join_aggr$created_2,"%m")
Join_aggr$created_3 <- format(Join_aggr$created_3,"%m")
Join_aggr$created_4 <- format(Join_aggr$created_4,"%m")
Join_aggr$created_5 <- format(Join_aggr$created_5,"%m")
Join_aggr$created_6 <- format(Join_aggr$created_6,"%m")
Join_aggr$created_7 <- format(Join_aggr$created_7,"%m")
Join_aggr$created_8 <- format(Join_aggr$created_8,"%m")
Join_aggr$created_9 <- format(Join_aggr$created_9,"%m")
Join_aggr$created_10<- format(Join_aggr$created_10,"%m")
Join_aggr$created_11<- format(Join_aggr$created_11,"%m")
Join_aggr$created_12<- format(Join_aggr$created_12,"%m")
Join_aggr$created_13<- format(Join_aggr$created_13,"%m")
Join_aggr$created_14<- format(Join_aggr$created_14,"%m")


#-----------------------------
Join_aggr$created_1 <- as.numeric(Join_aggr$created_1)
Join_aggr$created_2 <- as.numeric(Join_aggr$created_2)
Join_aggr$created_3 <- as.numeric(Join_aggr$created_3)
Join_aggr$created_4 <- as.numeric(Join_aggr$created_4)
Join_aggr$created_5 <- as.numeric(Join_aggr$created_5)
Join_aggr$created_6 <- as.numeric(Join_aggr$created_6)
Join_aggr$created_7<- as.numeric(Join_aggr$created_7)
Join_aggr$created_8 <- as.numeric(Join_aggr$created_8)
Join_aggr$created_9 <- as.numeric(Join_aggr$created_9)
Join_aggr$created_10 <- as.numeric(Join_aggr$created_10)
Join_aggr$created_11 <- as.numeric(Join_aggr$created_11)
Join_aggr$created_12 <- as.numeric(Join_aggr$created_12)
Join_aggr$created_13 <- as.numeric(Join_aggr$created_13)
Join_aggr$created_14 <- as.numeric(Join_aggr$created_14)


Join_aggr <- Join_aggr[is.na(Join_aggr)] = 0
# Join_aggr <- Join_aggr %>%
# mutate("Closed Date" = as.Date("Closed Date", format = "%m/%d/%y"))

```



```{r}
#x <- Join_aggr
#x[is.na(x)] = 0
 ggplot(x, aes(x=x$`Signed Date`, y=x$Additional.Amount_SUM, size = x$`Delay Days`)) +
     geom_point(alpha=0.8)

# x %>%
#   arrange(desc(x$`Signed Date`)) %>%
#   #mutate(x = factor(x$Number_OF_AA, x$Number_OF_AA)) %>%
#   ggplot(aes(x="Delay Days", y="Signed Date", size = "clint_age")) +
#     geom_point(alpha=0.5) +
#     scale_size(range = c(.1, 24), name="Population (M)")


x %>%   
  arrange(desc(x$Additional.Amount_SUM)) %>%
```


```{r}
x %>%
ggplot(aes(x=x$`Signed Date`, y=x$`Extensions per Loan`, size = x$`Delay Days`)) +
    #geom_point(alpha=0.8)
    geom_point(alpha=0.5, shape=21, color="blue") +
    scale_size(range = c(.1, 7), name="Population (M)") +
    scale_fill_viridis(discrete=TRUE, guide=FALSE, option="A") +
    theme_ipsum() +
    theme(legend.position="bottom") +
    ylab("Extension per Loan") +
    xlab("Signed Date") +
    theme(legend.position = "none")



x %>%
  arrange(desc(x$`Signed Date`)) %>%
  #mutate(country = factor(country, country)) %>%
  ggplot(aes(x=x$client_age, y=x$`Extensions per Loan`, size=x$Additional.Amount_SUM, x$`Main Amount`)) +
    geom_point(alpha=0.5, color="blue") +
    scale_size(range = c(.1, 5), name="Additional Amount_SUM (M)")
    theme(legend.position="bottom") 

plot(x$Additional.Amount_SUM, x$created_1)
```





```{r}
knitr::kable(
    rbind(
         c("-1.0 to -0.9" ,"Very strong", "Negative"),
         c("-0.9 to -0.7", "Strong", "Negative") ,
         c("-0.7 to -0.4", "Moderate", "Negative") ,
         c("-0.4 to -0.2", "Weak", "Negative"),
         c("-0.2 to 0","Negligible", "Negative") ,
         c("0 to 0.2","Negligible", "Positive"),
         c("0.2 to 0.4", "Weak", "Positive"), 
         c("0.4 to 0.7", "Moderate", "Positive"), 
        c("0.7 to 0.9", "Strong", "Positive"), 
         c("0.9 to 1.0", "Very strong", "Positive")), col.names=c("Correlation", "Strength", "Direction"), booktabs = TRUE)

cor(corr)

```

## Findings 

To summarize:
  -  Closed Date correlates (weak negative -0.314839425) with Extensions per Loan; </br>
  -  More loans will delay closed dates, which seems obvious; </br>
  -  the same status get Extensions per loan with Loan Closed Date -0.314839425;</br>
  - The signed dated and closed date shows the weak positive correlation; </br>
  - Delay Days and Closed Date -0.3793944729 weak negative correlation;</br>
  - Number_Of additional agreements have a strong positive correlation with    Additional.Amount_SUM, after each new agreement chance that sum will grow - also seems quite obvious </br>
  - Amount_SUM strong correlation with the first date where customer take an additional loan;</br>
  - Number_OF_AA   created_1  - 0.740886542; </br>
  - the rest of attr. doesn't provide info about significant or moderate corrlation. 
  
Interesting facts that the total sum of the additional agreement depends on the month where the first additional agreement has taken!!!
















